%
% ip.tex
%
% $Revision: 1.6 $
% $Date: 1992/07/21 00:12:54 $
%

\subsection*{NAME}

\noindent IP (Internet Protocol)

\subsection*{SPECIFICATION}

\noindent J. Postel. {\it Internet Protocol}. Request for Comments 
768, USC Information Sciences Institute, Marina del Ray, Calif., Aug. 1980.

\subsection*{SYNOPSIS}

\noindent IP handles fragmentation and routing required in
transmitting messages across heterogeneous interconnected networks.
This implementation is complete, with the exception of some of the
optional header fields.

\subsection*{REALM}

IP is in the ASYNC realm.

\subsection*{PARTICIPANTS}

IP removes a pointer to an IPhost
from the top of the stack of each participant.  If the local
participant is missing or if the local IPhost pointer is ANY\_HOST, 
IP will select an appropriate local IPhost.  

\subsection*{CONTROL OPERATIONS}

\begin{description}

\item[{\tt IP\_MYNET:}]
Return local host's IP network number.
This is an IP address with the host
component set to 0.  (session only)
\begin{description}
\item[{\rm Input:}] none
\item[{\rm Output:}] {\tt IPhost}
\end{description}

\item[{\tt IP\_REDIRECT:}] Modifies routing table to use a specified
gateway when delivering packets to a specified IP address.
The first address is for the destination
and the second is for the gateway.  (session or protocol)
\begin{description}
\item[{\rm Input:}] {\tt IPhost[2]}
\item[{\rm Output:}] none
\end{description}

\end{description}

\subsection*{CONFIGURATION}

IP must be configured above VNET:

\begin{verbatim}
         name=ip protocols=vnet;
\end{verbatim}

\medskip

If an explicit route for a remote network is not specified, 
IP will forward packets for that network to a {\em default gateway}.
The default gateway is initialized to IP\_SITE\_GTW from {\tt
xkernel/etc/site.h}, but this can be overridden by a rom file entry of the
form:
\begin{verbatim}
         ip    gateway     127.1.22.11
\end{verbatim}

If the default gateway is the distinguished address {\tt 0.0.0.0} or
can not be reached directly, IP will operate without a default gateway
and ERR\_XOBJ will be returned in cases where a default gateway would
otherwise have been used.

\subsection*{AUTHOR}

\noindent Clinton Jeffery, David Kays and Ed Menze
